---
title: "Statistical Report with Gemini Interpretation"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## Setup

Load required packages and configure your Gemini API key. Store your key in
an environment variable (`GEMINI_API_KEY`) rather than hard-coding it.

```{r}
#| label: setup

library(httr2)
library(glue)

# Reads your API key from the environment.
# Add this line to ~/.Renviron so it loads automatically:
#   GEMINI_API_KEY=AIzaSy-xxx
api_key <- Sys.getenv("GEMINI_API_KEY")
```

## Helper function: ask_gemini()

A small wrapper that sends a prompt to the Gemini API and returns the
reply as a plain character string.

```{r}
#| label: helper

ask_gemini <- function(prompt,
                       model   = "gemini-2.5-flash",
                       api_key = Sys.getenv("GEMINI_API_KEY")) {

  url <- glue(
    "https://generativelanguage.googleapis.com/v1beta/models/",
    "{model}:generateContent?key={api_key}"
  )

  request(url) |>
    req_body_json(list(
      contents = list(list(
        parts = list(list(text = prompt))
      ))
    )) |>
    req_retry(max_tries = 4, is_transient = \(r) resp_status(r) == 429) |>
    req_perform() |>
    resp_body_json() |>
    (\(r) r$candidates[[1]]$content$parts[[1]]$text)()
}
```

## Analysis

We run a small regression on the built-in `mtcars` dataset — fuel efficiency
(`mpg`) explained by engine displacement (`disp`) and horsepower (`hp`).

```{r}
#| label: analysis

model_fit <- lm(mpg ~ disp + hp, data = mtcars)
model_summary <- summary(model_fit)

# Capture the printed summary as text so we can pass it to Gemini
summary_text <- capture.output(print(model_summary)) |> paste(collapse = "\n")

cat(summary_text)
```

## Gemini Interpretation

We build a prompt that includes the raw R output and ask Gemini to
interpret the results in plain language.

```{r}
#| label: gemini-interpretation

prompt <- glue(
  "You are a statistical consultant writing for a non-technical audience. ",
  "Interpret the following R linear regression summary in 3–5 sentences. ",
  "Focus on practical meaning: which predictors matter, how well the model ",
  "fits, and any caveats worth noting.\n\n",
  "```\n{summary_text}\n```"
)

interpretation <- ask_gemini(prompt)
```

`r interpretation`

---

## Notes on workflow

* **`capture.output()`** converts any R printed output (summaries, tables,
  test results) into a character vector, making it easy to embed in a prompt.
* **`glue()`** keeps prompt construction readable; substitute `paste0()` if
  you prefer base R.
* **Model choice** — `gemini-2.5-flash` is fast and cheap for short
  interpretation tasks; swap in `gemini-2.5-pro` for more nuanced commentary.
* **Caching** — to avoid repeated API calls on every render, wrap
  `ask_gemini()` in `memoise::memoise()` or cache the result with
  `knitr::cache = TRUE` on the chunk.
* **API key safety** — never commit your key to version control. Use
  `.Renviron` or a secrets manager.
